Experiments with Synthetic Data

### 2022-07-23 dev/notebooks/synthetic.qmd
final streamlining

## Classifiers

```{julia}
using MLJ
N = 1000
X, ys = make_blobs(N, 2; centers=2, as_table=false, center_box=(-2 => 2), cluster_std=0.1)
ys .= ys.==2
X = X'
xs = Flux.unstack(X,2)
data = zip(xs,ys)
counterfactual_data = CounterfactualData(X,ys')
```

### Logistic Regression

```{julia}
using Flux
model = Chain(Dense(2,1))
M = FluxModel(model)
```

```{julia}
n_epochs = 10
using AlgorithmicRecourseDynamics.Models
M = Models.train(M,counterfactual_data; n_epochs=n_epochs)
```

```{julia}
plt_nn = plot(M,counterfactual_data;zoom=0)
savefig(plt_nn, joinpath(www_path, "nn_contour.png"))
```

![](../artifacts/upload/www/synthetic/nn_contour.png)

### Laplace Redux

```{julia}
λ = 0.1
model_laplace = Laplace(M.model, λ=λ)
LaplaceRedux.fit!(model_laplace, data)
```

```{julia}
M_laplace = Models.LaplaceModel(model_laplace)
```

```{julia}
plt_laplace = plot(M_laplace, counterfactual_data;zoom=0)
savefig(plt_laplace, joinpath(www_path, "la_contour.png"))
```

![](../artifacts/upload/www/synthetic/la_contour.png)

## Single Round

```{julia}
# Models:
models = (Bayesian=M_laplace, Plugin=M)
# Generators:
generators = (
    Greedy=GreedyGenerator(loss=:logitbinarycrossentropy), 
    Generic=GenericGenerator(loss=:logitbinarycrossentropy),
    REVISE=REVISEGenerator(loss=:logitbinarycrossentropy)
)
```

### Generate counterfactual

```{julia}
cb = false
γ = 0.75
mod = M_laplace
gen = generators[:Greedy]
plt_original = plot(mod,counterfactual_data;zoom=0,colorbar=false,title="(a)")
```

```{julia}
μ = 0.10
candidates = findall(ys.==0)
chosen = rand(candidates, Int(round(μ*length(candidates))))
X′ = copy(X)
y′ = copy(ys)
using CounterfactualExplanations.Counterfactuals: counterfactual, counterfactual_label
for i in chosen
    x = X[:,i]
    outcome = generate_counterfactual(x, 1, counterfactual_data, mod, gen,γ=γ)
    X′[:,i] = counterfactual(outcome)
    y′[i] = first(counterfactual_label(outcome))
end
counterfactual_data′ = CounterfactualData(X′,y′')
plt_single = plot(mod,counterfactual_data′;zoom=0,colorbar=false,title="(b)")
```

### Retrain

```{julia}
mod = Models.train(mod, counterfactual_data′; n_epochs=n_epochs)
plt_single_retrained = plot(mod,counterfactual_data′;zoom=0,colorbar=false,title="(c)")
```

### Repeat

```{julia}
i = 2
while i <= 10
    counterfactual_data′ = CounterfactualData(X′,y′')
    candidates = findall(y′.==0)
    chosen = rand(candidates, Int(round(μ*length(candidates))))
    H₀ = mod.model.H # use posterior as new prior
    mod = Models.train(mod, counterfactual_data′; n_epochs=n_epochs, H₀=H₀)
    for i in chosen
        x = X′[:,i]
        outcome = generate_counterfactual(x, 1, counterfactual_data, mod, gen; γ=γ)
        X′[:,i] = counterfactual(outcome)
        y′[i] = first(counterfactual_label(outcome))
    end
    i += 1
end
plt_single_repeat = plot(mod,counterfactual_data′;zoom=0,colorbar=false,title="(d)")
```

```{julia}
plt = plot(plt_original, plt_single, plt_single_retrained, plt_single_repeat, layout=(1,4), legend=false, axis=nothing, size=(600,165))
savefig(plt, joinpath(www_path, "poc.png"))
```

![](../artifacts/upload/www/synthetic/poc.png)

@@

```{julia}
using AlgorithmicRecourseDynamics.Experiments
using AlgorithmicRecourseDynamics.Experiments: Experiment
target = 1.0
experiment = Experiment(counterfactual_data, target, models, generators)
```

```{julia}
# Variables:
μ = 0.1
γ = 0.75
n_rounds = 2
n_folds = 5
T = 1000
```

```{julia}
output = run_experiment(experiment; n_rounds=n_rounds, μ=μ, n_folds=n_folds, γ=γ, T=T)
```

```{julia}
using CSV, BSON
CSV.write(joinpath(output_path,"synthetic_data.csv"), output)
BSON.@save joinpath(output_path,"synthetic_data.bson") output
```

>>>>

@@

### Data

```{julia}
catalogue = AlgorithmicRecourseDynamics.Data.load_synthetic()
```

```{julia}
using AlgorithmicRecourseDynamics.Models
using Flux
using AlgorithmicRecourseDynamics.Experiments
using AlgorithmicRecourseDynamics.Experiments: Experiment
using CSV, BSON

function _run_one()
    @info "Starting $(key)"

    # Models:
    model = Models.build_mlp()
    λ = 0.1
    model_laplace = Laplace(model, λ=λ)
    M = FluxModel(model)
    M_laplace = LaplaceReduxModel(model_laplace)
    models = (Bayesian=M_laplace, Plugin=M)

    # Pretrain:
    data_train, data_test = Models.train_test_split(counterfactual_data)
    models = map(model -> Models.train(model, data_train; n_epochs=100), models)

    # Generators:
    generators = (
        Greedy=GreedyGenerator(), 
        Generic=GenericGenerator(),
        REVISE=REVISEGenerator(),
        DICE=DiCEGenerator()
    )

    # Variables:
    μ = 0.1
    γ = 0.75
    n_rounds = 10
    n_folds = 5
    T = 1000
    target = 1.0
    experiment = Experiment(data_train, data_test, target, models, generators, 5)

    # Run:
    output = run_experiment(experiment; evaluate_every=2, n_rounds=n_rounds, μ=μ, n_folds=n_folds, γ=γ, T=T)

    CSV.write(joinpath(output_path,"output_$(key).csv"), output)
    BSON.@save joinpath(output_path,"output_$(key).bson") output
    BSON.@save joinpath(output_path,"experiment_$(key).bson") experiment

    @info "Successfully saved $(key)"

    return output, experiment
end

import AlgorithmicRecourseDynamics: run_experiment
function run_experiment(catalogue::Dict)
    output = [_run_one() for (key,counterfactual_data) in catalogue]
    return output
end
```

```{julia}
choices = [:linearly_separable, :overlapping, :plus_shaped, :balanced_positive_clusters]
data_sets = filter(p -> p[1] in choices, catalogue)
run_experiment(data_sets)
```

```{julia}
BSON.@load joinpath(output_path,"output_linearly_separable.bson") output
```

```{julia}
output[output.name.==:mmd_grid,:]
```
